Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-07-10 doc/manuscript/manuscript.Rmd
Suggested revisions and comments for methods.

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
The dry bean nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season.
Although the nursery locations were the same over sampling years, sampling plots within a location varied for sampling years.
Sampling of diseased tissue was carried out on a 'W' or 'V' shape transect, or randomly.
Because sampling depended on disease incidence, the number of samples isolated varied from year to year.
In total, we collected 366 isolates.

@@

Table: (\#tab:isolate-table) Description of *S. sclerotiorum* isolates used in this study. 
N = Number of Isolates. 
Key abbreviations: wmn = white mold nursery, unk = unknown.

Mycelial growth of each isolate was generated by activating sclerotia according to the method described previously [@matheron1989in-vitro].
These isolates were phenotypically characterized by determining mycelial compatibility group MCG and aggressiveness.
MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCG resulted in formation of a mycelial barrier [@kohn1990mycelial].
Aggressiveness of each isolate was assessed using a straw test that rates necrotic lesion on a scale from 1--9 with increasing aggressiveness [@petzoldt1996straw; @teran2006modified].
DNA was purified from mycelium of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.

@@

We characterized *S. sclerotiorum* isolates for microsatellite polymorphism using 16 primer pairs developed by @sirjusingh2001characterisation.
Polymerase chain reaction (PCR) was carried out as described using primers labeled with FAM fluorophore, and resulting PCR products were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range.
Capillary electrophoresis (fragment analysis) of PCR products added with size standard GeneScan™ 500 LIZ® was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center.

@@

@@

Of the 16 microsatellite loci, five included compound repeats, which made allele scoring difficult. 
These loci were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that enough loci have been sampled to sufficiently characterize the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
To control for non-independent samples (i.e. those sampled from the same location, region, year, and plant), data were clone-corrected for subsequent analyses.
The clone-corrected data set included 318 isolates.
All analyses were performed on this data set unless otherwise mentioned.
All data processing and analyses were performed in R version 3.4.0 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/

@@

A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* package version 2.4-3 [@vegan].
This method uses constrained ordinations on a distance matrix representing the response variable to delineate relative contribution of any number of independent explanatory variables. 
We used this method to delineate the phenotypic, geographic, and temporal components in influencing genetic composition of the populations.
<!-- RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables. -->
The distance matrix we used as the response variable was generated using Bruvo's genetic distance as implemented in *poppr* version 2.4.1, which employs a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr]. 
<!-- In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis. -->
To identify explanatory variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()` to identify the set of variables that explained more variation than predicted with the global model containing all variables.
An ANOVA was then performed to test for significance of the reduced model and marginal effects using 999 permutations.
The `varpart()` function of vegan was used to determine variation partitioning of explanatory variables.

@@

We used analysis of variance (ANOVA) to assess if MCGs differed in severity.
To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 57% of the isolates collected. Between-group differences were assessed using Tukey's HSD test using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].

@@

We wanted to assess if there was any correlation between multilocus genotypes (MLG) and mycelial compatibility groups (MCG). 
This was done using a graph-based approach with community analysis.
We first created a contingency table with MLGs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLG weighted by the number of samples shared in the connection. 
The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.0.1 was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
All communities that had fewer than 10 members were then consolidated into one.
These community definitions was then used to assess the average genetic distance (as defined by Bruvo's distance) within members of the community [@bruvo2004simple].

## Genetic Diversity {-}

@@

We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* version 2.4.1 to assess differentiation associated with wmn by quantifying the differentiation conferred by source and region [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
Source in this case was defined as a population belonging to a wmn or not and region defined as state. 
To identify specific regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version 2.1.0, assessing the per-sample posterior group assignment probability [@jombart2008adegenet]. 
This method decomposes the genetic data into principal components, and then uses a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant]. 
To avoid over-fitting the data, the optimal number of principal components was selected by using the *adegenet* function `xvalDapc()`. 
This function implements a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximizes the rate of successful group reassignment. 
We additionally assessed genotypic diversity by assessing Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis]. 
To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
Both genotypic diversity and linkage disequilibrium were calculated in *poppr* version 2.4.1 [@kamvar2014poppr].

@@

As many disparate geographic regions shared multilocus genotypes, we wanted to know if we could find any patterns in connectivity. 
To assess the connectivity of shared genotypes across regions, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
From these data, we were able to construct a graph with populations as nodes and shared genotypes as edges (connections) between the nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
Each node was weighted by the fraction of shared MLGs.
Each edge represents a single MLG, but because a single MLG could cross populations, that MLG may be represented by (n\*(n-1))/2 edges where n represents the number of populations it crosses.
Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
The graph was plotted using the R packages *ggplot2* version 2.2.1 [@ggplot2] and *ggraph* version 1.0.0 [@ggraph].
To ensure that we were able to capture the same community signal, we additionally ran this analysis including the 5 polymorphic markers described above.

@@

The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
The forward-backward selection process revealed severity, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the variation.
ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.032 (P = 0.001).
The marginal effects of state, host, and year were also significant (P = 0.001), but aggressiveness was not significant (P = 0.045).
Much of the variation appears to be driven by isolates from Mexico (Fig. \@ref(rda-biplot)) and 2005, but it's worth noting that all Mexican isolates were collected in 2005.
Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.

>>>>

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
> SEE: I was not sure what "dry bean nurseries" were because I'm use to seeing them referred to as the white mold nurseries.  I think if the introduction uses one term for it, this would be acceptable.  That or consider providing a description of the white mold nurseries in the methods section.  This now reminds me.. there is a publication by Lindsey Otto-Hanson and Steadman that gave a good description of these.  In fact, she called them multisite screening locations.  Here's the link: http://apsjournals.apsnet.org/doi/abs/10.1094/PDIS-11-10-0865

The dry bean nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season.
Although the nursery locations were the same over sampling years, sampling plots within a location varied for sampling years.
Sampling of diseased tissue was carried out on a 'W' or 'V' shape transect, or randomly.
Because sampling depended on disease incidence, the number of samples isolated varied from year to year.
In total, we collected 366 isolates that were used in the present study.

@@

Table: (\#tab:isolate-table) Description of *Sclerotinia sclerotiorum* isolates used in this study.
N = Number of Isolates. 
Key abbreviations: wmn = white mold nursery, unk = unknown.

Each isolate was regenerated by activating one or several sclerotia according to the method described previously [@matheron1989in-vitro].
Once actively growing, mycelial compatibility group (MCG) and aggressiveness was determined for each isolate.
MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCG resulted in formation of a mycelial barrier or barrage line [@kohn1990mycelial].
Since MCG are considered mutually exclusive, pairing was not required between all isolates in the study.
> SEE: Is that correct?  Perhaps consult with Becky or Rachana and then elaborate upon the sentence above.

Aggressiveness of each isolate was assessed using a straw test that rated necrotic lesion size on a scale from 1--9 with increasing aggressiveness [@petzoldt1996straw; @teran2006modified].
> SEE: Rating is not clear from the single sentence above.  On what plant host was it evaluated?  In the greenhouse?  What is a rating of 1 compared to a rating of 9 on the straw test?

DNA was purified from 100 mg mycelium collected from an actively growing culture of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
> SEE: More details on how tissue was disrupted may be needed here.

@@

We genotyped *S. sclerotiorum* isolates using 16 microsatellite primer pairs developed previously [@sirjusingh2001characterisation].
Polymerase chain reaction (PCR) was carried out as described using primers labeled with FAM fluorophore, and resulting PCR products were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range.
> SEE: Pretty rare to see PCR spelled out in full these days. I suggest just using the abbreviation and if PeerJ asks us to spell it out, we will.

Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center.

@@

> SEE: data were received as .fsa files, which are fasta-type files.  I have the raw data if you want them.  There is, apparently, an R-package called Fragman used for reading the files and can be used for scoring.  Anyway, in this case, I believe Rachana used PeakScanner (free software) to read the files and she recorded the fragment sizes manually (ie. read the chromatogram and wrote down in Excel what size the peak was).

@@

Of the 16 microsatellite loci, five included compound repeats, which made it difficult to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif. 
> SEE: I would not say allele scoring was difficult, because scoring an allele that is different than the size expected is really the same difficulty level as one that is at the expected size.  It's the accurate binning of alleles into the expected fragment size that is difficult.

Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
To control for non-independent samples (i.e. samples from the same location, region, year, and plant), data were clone-corrected for subsequent analyses.
> SEE: it is not clear what non-independent samples are.  Although your description is accurate and concise, a more general description might enable the reader to better understand what is meant here.

The clone-corrected data set included 318 isolates.
All analyses were performed on this data set unless otherwise mentioned.
All data processing and analyses were performed in R version 3.4.0 [@R] and are openly available and reproducible at https://github.com/everhartlab/sclerotinia-366/

@@

A distance-based redundancy analysis (dbRDA) [@legendre1999distance] was performed with the function `capscale()` in the *vegan* package version 2.4-3 [@vegan].
This method uses constrained ordinations on a distance matrix representing the response variable to delineate relative contribution of any number of independent explanatory variables. 
We used this method to delineate the phenotypic, geographic, and temporal components in influencing genetic composition of the populations.
<!-- RDA is a multivariate linear regression analysis of independent (explanatory: eg. environmental and phenotypic) variables on dependent (non-explanatory or response) variables. -->
The distance matrix we used as the response variable was generated using Bruvo's genetic distance as implemented in *poppr* version 2.4.1, which employed a stepwise mutation model for microsatellite data [@bruvo2004simple; @kamvar2014poppr]. 
<!-- In our case, a pairwise Bruvo's [@bruvo2004simple] distance matrix was generated from microsatellite data using *poppr* 2.4.1 [@kamvar2014poppr; @kamvar2015novel] and employed it as the dependent matrix for RDA analysis. -->
To identify explanatory variable(s) correlated with genetic variation, a forward-backward selection process was applied with the vegan function `ordistep()` to identify the set of variables that explained more variation than predicted with the global model containing all variables.
An ANOVA was then performed to test for significance of the reduced model and marginal effects using 999 permutations.
The `varpart()` function of vegan was used to determine variation partitioning of explanatory variables.

@@

We used analysis of variance (ANOVA) to assess if MCGs differed in aggressiveness.
To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 57% of the isolates collected. Between-group differences were assessed using Tukey's HSD test using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
> SEE: It's not clear what groups were used to compare aggressiveness in the ANOVA.  What alpha-level was used for the cut-off?

@@

We wanted to assess if there was correlation between multilocus genotypes (MLG) and mycelial compatibility groups (MCG). 
This was done using a graph-based approach with community analysis.
We first created a contingency table with MLGs and MCGs and converted it to a directed and weighted edgelist where each edge was a connection from an MCG to an MLG weighted by the number of samples shared in the connection. 
The cluster walktrap community detection algorithm as implemented in the `cluster_walktrap()` function in *igraph* version 1.0.1 was used to define clusters of MLGs and MCGs [@csardi2006igraph; @pons2006computing].
All communities that had fewer than 10 members were then consolidated into one.
These community definitions were then used to assess the average genetic distance (as defined by Bruvo's distance) within members of the community [@bruvo2004simple].

## Genetic Diversity {-}
> SEE: deleting this header?

@@

We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* version 2.4.1 to assess differentiation associated with wmn by quantifying the differentiation conferred by source and region [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
Source in this case was defined as a population belonging to a wmn or not and region defined as state. 
> SEE: is mtc included with "wmn"? At the beginning of the methods, I thought both were considered dry bean nurseries.

To identify specific regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version 2.1.0, assessing the per-sample posterior group assignment probability [@jombart2008adegenet]. 
This method decomposed the genetic data into principal components, and then used a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant]. 
To avoid over-fitting the data, the optimal number of principal components was selected by using the *adegenet* function `xvalDapc()`. 
This function implemented a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximized the rate of successful group reassignment. 
We additionally assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis]. 
To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
Both genotypic diversity and linkage disequilibrium were calculated in *poppr* version 2.4.1 [@kamvar2014poppr].

@@

As many disparate geographic regions shared multilocus genotypes, we wanted to know if we could find any patterns in connectivity. 
> SEE: avoid beginning a paragraph with a preposition because it forces the reader to read and remember the first half of the sentence before understanding it by reading the second half of the sentence.  It's better to reorganize and start with the subject, such as "Connectivity between shared multilocus genotypes identified in disparate geographic regions was evaluated."

To assess the connectivity of shared genotypes across regions, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
Each node was weighted by the fraction of shared MLGs.
Each edge represented a single MLG, but because a single MLG could cross populations, that MLG could be represented by (n\*(n-1))/2 edges where n represented the number of populations it crossed.
Edges were weighted by $1-P_{sex}$, where $P_{sex}$ was the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
The graph was plotted using the R packages *ggplot2* version 2.2.1 [@ggplot2] and *ggraph* version 1.0.0 [@ggraph].
To ensure that we captured the same community signal, we additionally performed this analysis including the 5 polymorphic markers described above.

@@

The dbRDA was performed on clone-corrected isolates (N = 318) after grouping them by location, host, region, and year.
The forward-backward selection process revealed aggressiveness, year, region, and host to be the optimal variables for the reduced model, accounting for 17.6% of the variation.
ANOVA revealed that the reduced model was significant with an adjusted R^2^ of 0.032 (P = 0.001).
The marginal effects of state, host, and year were also significant (P = 0.001), but aggressiveness was not significant (P = 0.045).
Much of the variation appears to be driven by isolates from Mexico (Fig. \@ref(rda-biplot)) and 2005, but it's worth noting that all Mexican isolates were collected in 2005.
Variance partitioning of the independent variables indicated aggressiveness to be the least influential factor with 0.1% contributing to explaining the variation of molecular data whereas the combination of variables accounted for 3.3%.
